---
title: "Machine Learning with `tidymodels`: Part 1" 
theme: readable
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 12
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(suppressWarnings({ 
  library(tidymodels)
  library(here)
  library(rio)
  library(vip)
}))
# prefer tidymodels function in any case of name conflict
tidymodels::tidymodels_prefer() 
```

## Introduction 

Welcome to Machine Learning with `tidymodels`. 

This workshop has three sections: 

1. We are going to discuss what machine learning is, what problems it works well for, and what problems it might work less well for. 
2. We will learn about the `tidymodels` framework to learn how to fit machine learning models in R. 
3. We will apply the `tidymodels` framework to explore multiple machine learning algorithms in R. 

## Resources 

The concepts contained herein draw on several fantastic resources that we highly recommend that you read and work through after the workshop. First we are indebted to [*Tidy Modeling with R* by Max Kuhn and Julia Silge](https://www.tmwr.org/). Second, much of the mathematical description comes from [*An Introduction to Statistical Learning* by James, Witten, Hastie, and Tibshirani](https://www.statlearning.com/) as well as ["Machine Learning Methods Economists Should Know About" by Athey and Imbens](https://arxiv.org/abs/1903.10075). 

Other useful free resources at varying levels of complexity are [*Feature Engineering and Selection: A Practical Approach for Predictive Models* by Kuhn and Johnson (2019)](http://www.feat.engineering/data-splitting.html), [*The Elements of Statistical Learning* by Hastie, Friedman, and Tibshirani](https://link.springer.com/book/10.1007/978-0-387-21606-5), and [*Deep Learning* by Goodfellow, Bengio, and Courville](https://www.deeplearningbook.org/)

## Required Packages

If you have not already done so, please run the following chunk to load all the packages needed for the course. 

```{r}
suppressMessages(suppressWarnings({ 
  library(tidymodels)
  library(here)
  library(rio)
  library(vip)
}))
# prefer tidymodels function in any case of name conflict
tidymodels::tidymodels_prefer() 
```

## Conceptualizing Machine Learning 

### What is Machine Learning? 

### What kind of problems does it work well on? 

### What kind of problems does it work less well on? 

### Supervised v. Unsupervised Learning 

This workshop focuses on supervised learning. 

## The `tidymodels` package

The `tidymodels` metapackage and framework is a collection of R packages for modeling that use tidyverse principles. `tidymodels` is designed like the [`tidyverse`](https://www.tidyverse.org/) as a set of modular packages, each with a narrow scope. Throughout the workshop we will highlight which packages `tidymodels` functions come from. A downside of a modular design is that there are lots of packages to keep track of as a user. Fortunately, the `tidymodels` metapackage loads a core set of `tidymodels` and `tidyverse` packages. 

A natural question is why does "tidiness" matter for machine learning? A strength of the R language is that developers can create any user-interface that fits their needs. Often, this leads to design for developers and not design for end users. Inconsistencies between naming conventions and arguments crop up between packages, which increases the time it takes to write code, and adds an unnecessary barrier for data analyses. 

In contrast, `tidymodels` follows a consistent set of design principles (dubbed [Design for Humans](https://design.tidyverse.org/unifying-principles.html)) applied for modeling code. For a new user, the main takeaways are that the defaults in `tidymodels` code are sensible, functions take the data structures that users have as opposed to the data structure that developers necessarily want, and there is a common syntax and structure to functions and arguments. 

As you advance in your machine learning journey, you may find that your needs are more specific than the ones provided in `tidymodels`. We hope that this introduction also gives you the appropriate terminology to reason through different packages and approaches. Throughout, we will highlight when a term is `tidymodels` specific or an application of a general concept. 

