---
title: "R Machine Learning: Introduction, Solutions"
theme: readable
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 12
    fig_height: 7
---

```{r setup, include=FALSE}
suppressMessages(suppressWarnings({ 
  library(tidymodels)
  library(palmerpenguins)
  library(ggplot2)
}))
# Prefer tidymodels functions in any case of name conflict
tidymodels::tidymodels_prefer() 
```

```{r}
penguins <- palmerpenguins::penguins
penguins
```

------------------------------------------------------------------------

### Challenge 1: Another Model

The function we seek to learn often has more than one variable. Write the equivalent mathematical function to illustrate the relationship between flipper length, bill depth, body mass, and island. Choose any feature and graph the relationship with a machine learning smoother.

```{r}
penguins %>%
  filter(!is.na(bill_depth_mm)) %>%
  ggplot(aes(x = body_mass_g, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```

------------------------------------------------------------------------

### Challenge 3: Classification and Regression

Explain whether the following are classification or regression problems.

1.  An advertiser is interested in the relationship between age and amount of hours of YouTube consumed. *Regression. Amount of hours is a continuous, quantitative variable.*

2.  A medical testing company conducts a procedure to determine whether a person has a cancer diagnosis. *cCassification. Either a person is diagnosed with cancer, or they are not.*

3.  A researcher is interested in the effect of an education intervention on students' test scores. *Regression. Test score is a quantitative variable.. However, we can imagine this problem being categorical if the test scores were not ordered. For example, we could classify someone's "learning style".*

------------------------------------------------------------------------
